---
title: "Legends"
author: "David"
toc: true
number-sections: true
format: 
  html: 
    embed-resources: true
---
## Basic Box Plot

- `library(ggplot2)`: Load the ggplot2 from library
- `ggplot(PlantGrowth, aes(x = group, y = weight))`:
  - Uses the `PlantGrowth` dataset
  - Creates a plot with `group` on the x-axis and `weight` on the y-axis
  - `aes()` defines the aesthetic mappings of the plot
- `geom_boxplot()`: Adds a box plot geometry to visualize the distribution of weights across different groups
- `ggtitle('Basic Box Plot')`: Add a title to the plot
- `theme(plot.title = element_text(hjust=0.5))`: Centers the title `(hjust=0.5)`

```{r}
library(ggplot2)
ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  labs(title = 'Basic Box Plot', 
       caption = 'David, DV, THU 2024') +
  theme(plot.title = element_text(hjust = 0.5))

```

## Add Legend

- `library(gridExtra)`: Load gridExtra from library
- `fill`: To fill the boxplot with colors and add a legend to inform the color
- `color`: To color the border of boxplot and add a legend to inform the color
- `grid.arrange()`: To arrange multiple boxplot
- `top`: To add a title in a arrange of boxplot
- `bottom`: To add a caption in a arrange of boxplot

```{r}
library(gridExtra)

d1 <- ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()
d2 <- ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_boxplot()
d3 <- ggplot(PlantGrowth, aes(x = group, y = weight, color = group)) + 
  geom_boxplot()

grid.arrange(d1, d2, d3, ncol = 3, top = 'Add Legends', bottom = 'David, DV, THU, 2024')
```
