---
title: "Legends"
author: "David"
toc: true
number-sections: true
format: 
  html: 
    embed-resources: true
---
## Basic Box Plot

- `library(ggplot2)`: Load the ggplot2 from library
- `ggplot(PlantGrowth, aes(x = group, y = weight))`:
  - Uses the `PlantGrowth` dataset
  - Creates a plot with `group` on the x-axis and `weight` on the y-axis
  - `aes()` defines the aesthetic mappings of the plot
- `geom_boxplot()`: Adds a box plot geometry to visualize the distribution of weights across different groups
- `ggtitle('Basic Box Plot')`: Add a title to the plot
- `theme(plot.title = element_text(hjust=0.5))`: Centers the title `(hjust=0.5)`

```{r}
library(ggplot2)
ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot() +
  labs(title = 'Basic Box Plot', 
       caption = 'David, DV, THU 2024') +
  theme(plot.title = element_text(hjust = 0.5))

```

## Add Legend

- `library(gridExtra)`: Load `gridExtra` from library
- `fill`: To fill the boxplot with colors and add a legend to inform the color
- `color`: To color the border of boxplot and add a legend to inform the color
- `grid.arrange()`: To arrange multiple boxplot
  - `ncol`: number input of column in one row
- `top`: To add a title in a arrange of boxplot
- `bottom`: To add a caption in a arrange of boxplot

```{r}
library(gridExtra)

d1 <- ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()
d2 <- ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_boxplot()
d3 <- ggplot(PlantGrowth, aes(x = group, y = weight, color = group)) + 
  geom_boxplot()

grid.arrange(d1, d2, d3, ncol = 3, top = 'Add Legends', bottom = 'David, DV, THU, 2024')
```

## Remove The Legend

`guides(fill = 'none')`: To remove the legend

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) + 
  geom_boxplot() +
  guides(fill = 'none') +
  labs(caption = 'David, DV, THU, 2024') +
  ggtitle('Remove Legend') +
  theme(plot.title = element_text(hjust = 0.5))
```

## Change The Color of The Legend

- `scale_fill_brewer(palette= "Pastel2")`: Change the color fill of the boxplot to `Pastel2`
- `ggtitle()`: Add a title to the boxplot

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  ggtitle('Change Color of Legend') +
  theme(plot.title = element_text(hjust=0.5)) +
  labs(caption = 'David, DV, THU, 2024')
```

## Change The Position of The Legend

`theme(legend.position = )`: To move the position of the legend
  - `top`: To move the legend to the top of the boxplot
  - `bottom`: To move the legend to the bottom of the boxplot
  - `right`: To move the legend to the right of the boxplot
  - `left`: To move the legend to the left of the boxplot

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  ggtitle('Change Color of Legend') +
  theme(plot.title = element_text(hjust=0.5),
        legend.position = 'top') +
  labs(caption = 'David, DV, THU, 2024')
```

## Practice

Practice changing color of the legend and change the position of the legend.

```{r}
d1 <- ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  ggtitle('Change Color of Legend') +
  theme(plot.title = element_text(hjust=0.5),
        legend.position = 'top')
d2 <- d1 + theme(legend.position = 'bottom')
d3 <- d1 + theme(legend.position = 'left')
d4 <- d1 + theme(legend.position = 'right')

grid.arrange(d1, d2, d3, d4, ncol=2,
             top='Chainging Position of Legend', bottom = 'David, DV, THU, 2024')
```